---
title: "Sample R Notebook"
author: "Christopher Prener, Ph.D."
date: '(`r format(Sys.time(), "%B %d, %Y")`)'
output: 
  github_document: default
  html_notebook: default 
---

## Introduction
This R Notebook models the submission process for assignments in all of my `R` based courses ([SOC 4930 & SOC 5050](https://slu-soc5050.github.io) and [SOC 4650 & SOC 5650](https://slu-soc5650.github.io)). Make sure your notebooks are formatted like this one before you submit your work!

## Project Set Up
Your notebook should begin with an introduction and then a "Project Set Up" section that includes the following code chunk:

```{r setup}
knitr::opts_knit$set(root.dir = here::here())
```

Since our notebook resides in a subdirectory of the R project called `doc`, all relative paths to folders as in the `import("data/sampleData.csv")` are assumed to subdirectories of `doc`. This, of course, is not the case. Rather, we want our notebook to assume that relative paths of the folder with the R project file in it. The `knitr` and `here` package functions in the `setup` chunk work together to make sure that your notebook is compatible with the [project directory structure](https://chris-prener.github.io/SSDSBook/organizing-projects.html) we'll be using this semester.

## General Notebook Structure
Up above the introduction is a [`YAML`](https://en.wikipedia.org/wiki/YAML) header that contains, by default, fields for `title` and `output`. Make sure you change the `output` from only saying `html_notebook` to also including `github_document` for each assignment. Both output types should be noted as `default`. Also be sure to add in and fill out the `author` and `date` fields. The `date` field function can be inserted into every document without change. When your notebook is "knit", it will always insert the current date. 

I recommend using second level headings (`##`) for major sections, such as the different parts of each assignment. For each question, use a third level heading (`###`).

Question responses show follow a general "literate programming" pattern below each heading where you:

1. Describe how the code addresses the question
2. Execute the code in a code chunk, which should be named to keep your notebook's output organized and easy to look through if necessary
3. Provide a written interpretation of the results

Your narrative text (1 and 3) should be written in full sentances.

## Dependencies
Your notebooks should always have a clearly organized dependencies section after the project set up code chunk. Be sure to include all of the packages that your document requires. This document requires the following packages:

```{r dependencies}
library(dplyr)    # data wrangling
library(ggplot2)  # plotting

library(skimr)    # descriptive statistics
```

I recommend adding a note next to each package that specifies why you are including it. This will help you in the long run, but making sure you have a clear sense of what each package does and helping build "muscle memory" of these functions by typing it out repeatedly. I also recommend grouping packages by type. Above, the two `tidyverse` packages are placed in the code chunk together.

## Part 1
### Question 1
When you execute code within the notebook, the results appear beneath the code. Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Cmd+Shift+Enter*. 

```{r q1-plot}
ggplot(data = mpg, mapping = aes(hwy)) +
  geom_histogram(bins = 28)
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I* (on an Apple computer).

If questions ask you to create "publication ready" plots, make sure you save them to the results directory of your assignment's project:

```{r q1-plotSave}
plot <- ggplot(data = mpg, mapping = aes(hwy)) +
          geom_histogram(bins = 28) +
          labs(
            title = "Histogram of Highway Fuel Efficiency",
            xlab = "highway fuel efficiency (miles per gallon)"
          )
ggsave("results/q1.png", plot)
```

### Question 2
Question two asks for descritive statistics, which we can obtain using the `skimr()` function:

```{r q2}
skim(mpg, hwy, cty)
```

The average highway fuel efficiency for the vehicles in the sample is 23.44 miles per gallon while the average city fueld efficiency is only 16.86 miles per gallon.